---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load the data.

```{r, echo=TRUE}
data <- read.csv(unz("activity.zip", "activity.csv"), header=T, quote="\"", sep=",",colClasses=c("integer","Date","integer"))
```


## What is mean total number of steps taken per day?

Here is a histogram of total number of steps taken per day
```{r, echo=TRUE,fig.height=4}
totalStepsPerDay <- aggregate(steps~date,data=data,FUN=sum)
hist(totalStepsPerDay$steps,col=c("red"),xlab="Total number of Steps per Day")
```

Calculating the mean and median total number of steps taken per day.
```{r, echo=TRUE}
m1 <- mean(totalStepsPerDay[,2])
m2 <- median(totalStepsPerDay[,2])
```
The mean and median total number of steps taken per day are `r m1` and `r m2` respectively.

## What is the average daily activity pattern?

This is a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.
```{r, echo=TRUE,fig.height=4}
avgDailyData <- aggregate(steps~interval,data=data,FUN=mean)
library(lattice)
xyplot(steps ~ interval , data = avgDailyData, type="l",xlab="Interval",ylab="Number of steps")

  
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
avgDailyData[avgDailyData$steps==max(avgDailyData$step),"interval"]
```

## Inputing missing values

Here is the total number of missing values in the dataset (i.e. the total number of rows with NAs).
```{r, echo=TRUE}
nrow(data[is.na(data$steps),])
```

Fill in the missing values using mean for that day.
```{r, echo=TRUE}
# calculate mean for each 5 minute interval
library(plyr)
meanStepsByInterval <- ddply(data,.(interval),summarize,mean=mean(steps,na.rm=TRUE))
# replace NA values
dataComplete <- data
j=0
for (i in 1:nrow(dataComplete)) {
  if (is.na(dataComplete[i,"steps"])) {
    j=j+1
    dataComplete[i,"steps"] <- meanStepsByInterval[meanStepsByInterval$interval==dataComplete[i,"interval"],"mean"]
  }
}
print(j)
```

Here is a histogram of total number of steps taken per day  
```{r, echo=TRUE,fig.height=4}
totalStepsPerDay <- aggregate(steps~date,data=dataComplete,FUN=sum)
hist(totalStepsPerDay$steps,col=c("blue"),xlab="Total number of Steps per Day")
```

Calculating the mean and median total number of steps taken per day.
```{r, echo=TRUE}
m1 <- mean(totalStepsPerDay[,2])
m2 <- median(totalStepsPerDay[,2])
```
The mean and median total number of steps taken per day are `r m1` and `r m2` respectively.

As can be seen, the impact of filling in NA data is (1) no change to the mean(2) the median has become the same with the mean.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with the filled-in missing values with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
l <- levels(as.factor(data$date))
w <- sapply(l,function(x) ifelse(weekdays(as.Date(x)) %in% c("Saturday","Sunday"),"weekend","weekday"))
dataComplete <- mutate(dataComplete, typeOfDay=w[as.character(date)])
```

This is a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r, echo=TRUE}
avgDailyData <- aggregate(steps~interval+typeOfDay,data=dataComplete,FUN=mean)
library(lattice)
xyplot(steps ~ interval | typeOfDay, data = avgDailyData, type="l", xlab="Interval",ylab="Number of steps", layout = c(1, 2))

```

As can be seeen from above, in general there is higher number of average steps during weekend compared to weekday for each 5-min interval during the  active intervals.